\tit{Waarmee} vs. \tit{met wat}\\ Constituency spellout, r-pronouns

### 2020-01-11 paper.tex
only formatting
left to do
-move tables to end
-double + single spacing
-abstract
-personal details
-keywords
weg bij jan vandaan
eigenlijk alleen koreaan verwijderd en ossetic erbij
nu op naar prefix vs. suffix!

\documentclass{article}

@@

\title{\tit{Waarmee} vs. \tit{met wat}}
\author{Fenna Bergsma}
\date{\today}

@@

@@

This section showed that being a proper constituent is a necessary requirement for \tit{waarmee} `with what' to surface. The next section elaborates on how proper constituency and spellout relate to each other.

\section{\tit{Waarmee} `with what' as suppletion}\label{sec:suppletion}

@@

In this section I start with showing a similar instance in which proper constituency and spellout go hand in hand. Then I show how these facts can be straightforwardly follow from the spellout algorithm in Nanosyntax. The section ends with showing how this logic can be applied to \tit{waarmee} `with what' and \tit{met wat} `with what'.

In what follows I show another instance in which proper constituency influences spellout, namely suppletive negative forms in Korean, as discussed by \citet{chung2007}. The point is that suppletion only goes through when the relevant features form a proper constituent, and not when an additional structure intervenes.

Sentences in Korean can be negated by the two negative prefixes \tit{ani} or \tit{mos}, as shown in \ref{ex:koreannegation}.

\ex.\label{ex:koreannegation}
\ag. ca -n -ta\\
 sleep -\tsc{pres} -\tsc{decl}\\
 `is sleeping'
\bg. mos/an(i) ca -n -ta\\
 \tsc{neg} sleep -\tsc{pres} -\tsc{decl}\\
 `cannot sleep / is not sleeping' \hfill \citep{chung2007}

The verb \tit{al-} `know' does not combine with any of the negative markers, but it it uses the suppletive stem \tit{molu.\tsc{neg}}. This is illustrated in \ref{ex:koreanknow}.

\ex.\label{ex:koreanknow}
\ag. al -n -ta\\
know \tsc{-pres} \tsc{-decl}\\
`know(s)'
\bg. *mos/*an(i) al -n -ta\\
\tsc{neg} know -\tsc{pres} -decl\tsc{}\\
\bg. molu -n -ta\\
\tsc{neg}.know -\tsc{pres} -\tsc{decl}\\
`do(es) not / cannot know'\label{ex:molu} \hfill \citep{chung2007}

When \tit{al-} `know' is causativized before it is negated, the regular stem is used again, shown in \ref{ex:knowcausneg}. \ref{ex:molucaus} illustrates that the suppletive stem \tit{molu-} is no longer grammatical.

\ex.
\ag. al -li-\\
 know -\tsc{caus}\\
 `let know, inform'
\bg. ani/ mos al -li -ess -ta\\
 \tsc{neg} \tsc{neg} know -\tsc{caus} -\tsc{past} -\tsc{decl}\\
 `did not /could not inform'\label{ex:knowcausneg}
\bg. *molu -li -ess -ta\\
 \tsc{neg}.know -\tsc{caus} -\tsc{past} -\tsc{decl}\\\label{ex:molucaus} \hfill \citep{chung2007}

\citet{chung2007} points out that it cannot be phonological adjacency of the negation and \tit{al-} `know' that causes the suppletion in \ref{ex:molu}. The reason for that is that in \ref{ex:knowcausneg} \tit{al-} also follows the negation, but there is no suppletion. \citet{chung2007} suggests crucial difference has to do with in proper constituency.
The structure of \ref{ex:molu} is given in \ref{ex:structuremolu}: Neg and \tit{al-} `know' form a proper constituent. \ref{ex:structurenegmolu} represents the structure of \ref{ex:knowcausneg}. Neg and \tit{al-} `know' no longer form a proper constituent, because \tsc{caus} is also part of the structure as the sister of \tit{al-} `know'.

\ex.
%%%
\a. \begin{forest} boom
[
    [Neg]
    [know]
]
\end{forest}\label{ex:structuremolu}
\b. \begin{forest} boom
[
    [Neg]
    [
        [know]
        [\tsc{caus}]
    ]
]
\end{forest}\label{ex:structurenegmolu}

\citet{caha2009} uses this pattern to argue in favor of a model with phrasal spellout rather than inserting phonology into terminal nodes, specifically Nanosyntax \citep{starke2009}. The idea is that the lexicon contains syntactic structures that are linked to a particular piece of phonology. If Korean has a lexical entry as in \ref{ex:entrymolu}, this can be inserted into the syntactic tree in \ref{ex:structuremolu}.

\ex. \begin{forest} boom
[
    [Neg]
    [know]
]
  {\draw (.east) node[right]{⇔ \tit{molu}}; }
\end{forest}\label{ex:entrymolu}

Since spellout targets constituent \citep{starke2009}, \ref{ex:entrymolu} is not inserted into the structure in \ref{ex:structurenegmolu}. \tsc{caus} is contained in the syntactic structure but not in the structure in the lexicon in \ref{ex:entrymolu}.

The same logic can be applied to \tit{waarmee} `with what' and \tit{met wat} `with what'. The preposition and wh-element together form a proper constituent, as in \ref{ex:structurewaarmee}, or the embedded CP of the relative clause is a sister of the wh-element, as in \ref{ex:structuremetwat}.

\ex.
\a. \begin{forest} boom
[
    [P]
    [Wh]
]
\end{forest}\label{ex:structurewaarmee}
%%
\b. \begin{forest} boom
[
    [P]
    [
        [Wh]
        [CP
            [...,roof]
        ]
    ]
]
\end{forest}\label{ex:structuremetwat}

Assume we have a lexical entry for \tit{waarmee} `met wat' that look as \ref{ex:entrywaarmee}. This entry can only be inserted into \ref{ex:structurewaarmee}, because the lexical structure matches the syntactic structure. It cannot be into \ref{ex:structuremetwat}, because the lexical structure does not contain the CP.

\ex. \begin{forest} boom
[
    [InsP]
    [WhP]
]
  {\draw (.east) node[right]{⇔ \tit{waarmee}}; }
\end{forest}\label{ex:entrywaarmee}

What I have shown in this section is how proper constituency relates to a change in phonology. However, this analysis missed numerous phonological properties of \tit{waarmee} `with what' and \tit{met wat} `with what', e.g. that they both contain the morpheme \tit{wa}. Under the suppletion analysis above, one can the phonological form of \tit{waarmee} could just as well be completely phonologically unrelated. The point of presenting the problem like this is to illustrate that the distribution between \tit{waarmee} `with what' and \tit{met wat} `with what' can be captured as a matter of spellout.

@@

\section{A finer decomposition}\label{sec:finer}

In this section I investigate the internal structure of \tit{waarmee} `with what' and \tit{met wat} `with what' to capture the phonological similarities between the two forms. First, I identify \tit{w} and \tit{a} as morphemes that appear in both expressions. Putting these two aside, I concentrate on \tit{ermee} `with it' and \tit{met 't} `with it'. I decompose these two expressions further, and end up arguing for a picture as shown in Table \ref{tbl:nofeatures}.
The elements \tit{w} and \tit{a} express the same syntactic structure in both expressions. The elements \tit{met} and \tit{'t} together also express the same features as \tit{mee} and \tit{er} together, but the distribution differs. \tit{Met} expresses less structure than \tit{mee} and \tit{'t} expresses more structure than \tit{er}. It becomes clear that the \tit{er} in fact corresponds to the locative \tit{r} in Dutch.

@@

Finally, I put everything together and show how \tit{waarmee} `with what' is inserted if all features form a constituent, and \tit{met wat} `with what' if this is not the case.

\subsection{Taking everything apart}

In this section I decompose \tit{waarmee} `with what' and \tit{met wat} `with what'. The idea is that both expressions realize the same features, and \tit{waarmee} `with what' takes precedence when all features form a constituent. The morphemes I distinguish in \tit{waarmee} `with what' are \tit{w}, \tit{-aa-}, \tit{-r} and \tit{mee}. Within \tit{met wat} `with what' I distinguish \tit{met}, \tit{w}, \tit{a} and \tit{-t}. \ref{ex:decompose} shows this as well.

@@

@@

\subsection{Putting it back together}

>>>>

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
%\usepackage{setspace}
%\doublespacing

@@

\title{\tit{Waarmee} vs. \tit{met wat}\\
Constituency spellout, r-pronouns}
\author{Fenna Bergsma}
\date{\today}

@@

\raggedright

@@

@@

@@

\section{Taking \tit{waarmee} and \tit{met wat} apart}

In this section I investigate the internal structure of \tit{waarmee} `with what' and \tit{met wat} `with what' to capture the phonological similarities and differences between the two forms. First, I identify \tit{w} and \tit{a} as morphemes that appear in both expressions. Putting these two aside, I concentrate on \tit{ermee} `with it' and \tit{met 't} `with it'. I decompose these two expressions further, and end up arguing for a picture as shown in Table \ref{tbl:nofeatures}.
The elements \tit{w} and \tit{a} express the same syntactic structure in both expressions. The elements \tit{met} and \tit{'t} together also express the same features as \tit{mee} and \tit{er} together, but the distribution differs. \tit{Met} expresses less structure than \tit{mee}, and \tit{'t} expresses more structure than \tit{er}. \tit{Met} is shown to be an prefix, and \tit{mee} is a suffix. It also becomes clear that the \tit{er} in fact corresponds to the locative \tit{r} in Dutch.

@@

 The idea is that both expressions realize the same features, and \tit{waarmee} `with what' takes precedence when all features form a constituent. The morphemes I distinguish in \tit{waarmee} `with what' are \tit{w}, \tit{-aa-}, \tit{-r} and \tit{mee}. Within \tit{met wat} `with what' I distinguish \tit{met}, \tit{w}, \tit{a} and \tit{-t}. \ref{ex:decompose} shows this as well.

@@

\section{Two mechanisms}

\subsection{Suppletive obliques}

\begin{table}[ht]
	\center
	\caption {Iron Ossetic pronoun and noun}
	\begin{minipage}{0.56\linewidth}
		\begin{tabularx}{\textwidth}{ccccccc}
		\toprule
              & 1.\tsc{sg}  & head    \\
		\midrule
    \tsc{nom} & æz          & sær-Ø   \\
    \tsc{acc} & mæn-Ø       & sær-Ø   \\
    \tsc{gen} & mæn (⁇?)    & sær-y   \\
    \tsc{ins} & mæn-æj      & sær-æj  \\
    \tsc{dat} & mæn-æn      & sær-æn  \\
    \tsc{all} & mæn-mæ      & sær-mæ  \\
    \tsc{ade} & mæn-yl      & sær-yl  \\
    \tsc{eq}  & mæn-au      & sær-au  \\
    \tsc{com} & mæn-imæ     & sær-imæ \\
    \bottomrule
\end{tabularx}
\label{tbl:ossetic}
\end{minipage}
\end{table}

now the question: is \tit{æz} a suppletive nominative or is \tit{mæn} the suppletive stem from the accusative on? ossetic has a way to tell: suspended affixation.

What we see in (3-a) is an ordinary coordination of two phrases, each marked
by a plural marker and a case marker. Suspended affixation is then illustrated in
(3-b). It is a type of a coordination where the comitative case marker is located
only on the second conjunct, with no change in interpretation. The first conjunct
simply bears no case marker in such cases.

\ex.
\ag. bæx-t-imæ æmæ gæl-t-imæ\\
horse-pl-com and ox-pl-com\\
\bg. bæx-tæ æmæ gæl-t-imæ\\
horse-pl and ox-pl-com\\
`with horses and oxen' \hfill (Iron Ossetic, Erschler 2012: 165)

doing this with 1\tsc{sg} shows that the bare stem is \tit{mæn}.

\ex.
\ag. mæn æmæ Zauyr-æn\\
 1.\tsc{sg} and Zaur-\tsc{dat}\\
\bg. *æz æmæ Zauyr-æn\\
 1.\tbf{sg} and Zaur-\tsc{dat}\\
 `me and Zaur' \hfill (Belyaev 2014: 39)

\tit{mæn} is actually the caseless form, \tit{æz} is marked for nominative.

my conclusion: Dutch has a suppletive form for nominative, accusative and dative. Then higher cases do not combine with that suppletive form, but with the unmarked form.

't
't
't
er-van
er-mee
er-mee

\subsection{Prefix or suffix}

https://ling.auf.net/lingbuzz/001026

@@

\section{Showing it in a derivation}
